---
title: "`r paste(params$species_proper, params$era, params$area_name)` bottom trawl CPUE"
author: "`r params$author`"
date: "`r Sys.Date()`"
output: html_document
params:
   author: "Sean C. Anderson"
   species_proper: "Pacific Cod"
   area: !r c("3[CD]+")
   area_name: !r c("3CD")
   skip_single_variable_models: TRUE
   era: "modern"
   april1_year: FALSE
   min_year_historic: 1956
---

```{r setup, include=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(params$species_proper)))
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.asp = 0.618,
  cache.path = paste0(spp, "-", params$era, "-", params$area_name, "-cache/"),
  fig.path = paste0(spp, "-", params$era,  "-", params$area_name, "-fig/"),
  echo = TRUE,
  autodep = TRUE,
  cache = TRUE,
  cache.comments = FALSE
)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library("dplyr")
library("ggplot2")
library("gfplot")
library("here")
library("readr")
library("purrr")
library("glmmTMB")
ggplot2::theme_set(gfplot::theme_pbs())
```

```{r cpue-params-print}
print(params)
```

```{r fishing_event_trip, echo=FALSE}
fishing_event_trip <- if (params$era == "modern") "fishing event" else "trip"
```

```{r cpue}
if (params$era == "modern") {
  fi <- here("data/cpue-modern.rds")
  if (!file.exists(fi)) {
    d1996 <- gfplot::get_cpue_index(gear = "bottom trawl", min_cpue_year = 1996)
    write_rds(d1996, fi)
  } else {
    d1996 <- read_rds(fi)
  }
} else {
  fi <- here("data/cpue-historic.rds")
  if (!file.exists(fi)) {
    d <- gfplot::get_cpue_historic(species = NULL, end_year = 1995, 
      alt_year_start_date = "04-01")
    write_rds(d, fi)
  } else {
    d <- read_rds(fi)
  }
}
```

Define our fleet. See the help for `?gfplot::tidy_cpue_index` and `?gfplot::tidy_cpue_historic` to see a definition of these arguments. <https://github.com/pbs-assess/gfplot>

```{r define-fleet}
if (params$era == "modern") {
  define_fleet <- function(area, area_name) {
    out <- gfplot::tidy_cpue_index(d1996,
      species_common = tolower(params$species_proper),
      gear = "bottom trawl",
      alt_year_start_date = "04-01",
      use_alt_year = params$april1_year,
      year_range = c(1996, 2017), 
      lat_range = c(48, Inf),
      min_positive_tows = 100,
      min_positive_trips = 5,
      min_yrs_with_trips = 5,
      depth_band_width = 25,
      area_grep_pattern = area,
      depth_bin_quantiles = c(0.001, 0.999),
      min_bin_prop = 0.001,
      lat_band_width = 0.1)
    out$area <- area_name
    out
  }
  dfleet <- map2(params$area, params$area_name, define_fleet)
} else {
  define_fleet <- function(area, area_name) {
    out <- gfplot::tidy_cpue_historic(d,
      species_common = tolower(params$species_proper),
      use_alt_year = params$april1_year,
      year_range = c(params$min_year_historic, 1995),
      depth_band_width = 25,
      area_grep_pattern = area,
      depth_bin_quantiles = c(0.001, 0.999),
      min_bin_prop = 0.001)
    out$area <- area_name
    out
  }
  dfleet <- map2(params$area, params$area_name, define_fleet)
}
```

The most frequent factor levels for `r fishing_event_trip`s that caught `r params$species_proper` are the reference levels. This only affects the relative scale of the final CPUE index, not the shape, since there are no fixed-effect interactions in our standardization model.

Here, and throughout, `pos_catch` refers to if a `r fishing_event_trip` caught (was positive) for `r params$species_proper`.

Total `r params$species_proper` catch and total effort for our 'fleet':

```{r catch-effort, fig.asp=0.9}
dfleet %>% bind_rows() %>% 
  group_by(year, area) %>% 
  summarise(
    `Species catch` = sum(spp_catch)/1000, 
    `Hours fished` = sum(hours_fished)/1000) %>% 
  reshape2::melt(id.vars = c("year", "area")) %>%
  ggplot(aes(year, value)) +
  geom_line() +
  facet_grid(variable~area, scales = "free_y") +
  ylab("Value (1000 kg or 1000 hours)") + xlab("") +
  ylim(0, NA)
```

Look at depth distribution across all `r fishing_event_trip`s:

```{r cpue-depth, fig.asp=0.618}
depth_bands <- as.numeric(as.character(unique(bind_rows(dfleet)$depth)))
dfleet %>%
  bind_rows() %>% 
  mutate(`Trip or fishing event\ncaught this species` = 
      ifelse(pos_catch == 1, "Yes", "No")) %>% 
  ggplot(aes(best_depth, fill = `Trip or fishing event\ncaught this species`)) +
  geom_histogram(binwidth = 10) +
  ylim(0, NA) +
  geom_vline(xintercept = depth_bands, lty = 2, col = "grey80") +
  coord_cartesian(expand = FALSE) +
  facet_wrap(~area, ncol = 2)
```

Make bubble plots of `r fishing_event_trip` counts. The open circles represent all `r fishing_event_trip`s and the shaded circles represent `r fishing_event_trip`s that were positive for `r params$species_proper`.

```{r bubble-plots, fig.asp=0.75, echo=FALSE}
group <- if (params$era == "modern") "fishing_event_id" else "trip_id"
for (i in seq_along(dfleet)) {
  gfplot:::plot_predictor_bubbles(dfleet[[i]], "locality",
    group = group) %>% print()
  
  dfleet[[i]] %>% mutate(depth = as.factor(as.character(depth))) %>% 
    gfplot:::plot_predictor_bubbles("depth",
      group = group) %>%
    print()
  
  if (params$era == "modern") {
    dfleet[[i]] %>% mutate(latitude = as.factor(as.character(latitude))) %>% 
      gfplot:::plot_predictor_bubbles("latitude",
        group = group) %>%
      print()
    
      
  gfplot:::plot_predictor_bubbles(dfleet[[i]], "vessel",
    group = group) %>%
    print()
  }
  
  dfleet[[i]] %>% mutate(month = as.factor(as.character(month))) %>% 
    gfplot:::plot_predictor_bubbles("month",
      group = group) %>%
    print()
}
```

Now fit the models. These are GMLMs with a Tweedie / compound-Poisson-Gamma observation model. Note that the `p` parameter, in [TMB notation](https://kaskr.github.io/adcomp/group__R__style__distribution.html#ga262f3c2d1cf36f322a62d902a608aae0), is bounded to be between 1 (fully Poisson) and 2 (fully Gamma). We will fit a model with only the year predictor`r if (params$skip_single_variable_models) "" else ", a model with each of the other predictors and year,"` and then a model with all of the predictors. The full model in lme4 syntax is: 
`r if (params$era == "modern") "cpue ~ 0 + year_factor + depth + month + latitude + (1 | locality) + (1 | vessel) + (1 | year_locality)" else "cpue ~ 0 + year_factor + depth + month + (1 | locality) + (1 | year_locality)"`.

```{r fit-cpue, results='hide', warning=FALSE, message=FALSE}
for (i in seq_along(dfleet)) {
  dfleet[[i]]$year_locality <- paste(dfleet[[i]]$year_factor, dfleet[[i]]$locality)
}

if (params$era == "modern") {
  formulas <- data_frame(
    formula = c(
      "cpue ~ 0 + year_factor",
      "cpue ~ 0 + year_factor + depth",
      "cpue ~ 0 + year_factor + month",
      "cpue ~ 0 + year_factor + latitude",
      "cpue ~ 0 + year_factor + (1 | locality)",
      "cpue ~ 0 + year_factor + (1 | vessel)",
      "cpue ~ 0 + year_factor + depth + month + latitude + (1 | locality) + (1 | vessel) + (1 | year_locality)"
    ),
    formula_version = c(
      "Unstandardized",
      "Depth",
      "Month",
      "Latitude",
      "Locality",
      "Vessel",
      "Full standardization"
    )
  )
} else {
  formulas <- data_frame(
  formula = c(
    "cpue ~ 0 + year_factor",
    "cpue ~ 0 + year_factor + depth",
    "cpue ~ 0 + year_factor + month",
    "cpue ~ 0 + year_factor + (1 | locality)",
    "cpue ~ 0 + year_factor + depth + month + (1 | locality) + (1 | year_locality)"
  ),
  formula_version = c(
    "Unstandardized",
    "Depth",
    "Month",
    "Locality",
    "Full standardization"
  )
)
}

torun <- expand.grid(formula = formulas$formula,
  area = params$area_name, stringsAsFactors = FALSE)
torun <- inner_join(torun, formulas, by = "formula")

if (params$skip_single_variable_models) {
  torun <- filter(torun, 
    formula_version %in% c("Unstandardized", "Full standardization"))
}

file_model <- here(paste0("data/generated/cpue-models-", spp, "-", params$era,  
  "-", params$area_name,  ".rds"))
if (!file.exists(file_model)) {
  system.time({
  model <- plyr::mlply(torun, function(formula, area, formula_version) {
    df <- dfleet[[which(params$area_name == area)]]
    message("Fitting area ", area, " and model ", formula)
    fit_cpue_index_glmmtmb(df, as.formula(formula))
    })
  })
  saveRDS(model, file_model)
} else {
  model <- readRDS(file_model)
}

predictions <- plyr::ldply(model, predict_cpue_index_tweedie)
write_csv(predictions, 
  here(paste0("data/generated/cpue-predictions-", spp, "-", params$era, "-", params$area_name, ".csv")))
```

In the predictions data frame (file `r paste0("cpue-predictions-", spp, "-", params$era,  "-", params$area_name, ".csv")`, `est_link` is the estimate on the link (log) scale, `se_link` is the standard error on the link scale, and `est`, `lwr`, `upr` are the estimate, lower 95%, and upper 95% confidence intervals on the natural (exponentiated) scale in kg/hours fished. All of the model versions are included.

```{r header-show, echo=FALSE, eval=FALSE}
glimpse(predictions)
```

Now we will plot the coefficients from the full standardization model. Note that the estimates are with respect to the base or reference level, which has been set to the most common level for positive `r fishing_event_trip`s. The reference level is omitted from each of the panels.

```{r coef-plot-funcs, fig.asp=1.4, fig.width=8, echo=FALSE}
make_fe_plots <- function(object) {
  su <- summary(object)$coefficients$cond
  sud <- as.data.frame(su)
  sud$param <- row.names(su)
  row.names(sud) <- NULL
  sud <- rename(sud, est = Estimate, se = `Std. Error`)
  sud <- mutate(sud, par_value = gsub("^[A-Z_a-z]+", "", param))
  sud <- mutate(sud, par_group = gsub("^([A-Z_a-z]+)[0-9.]+$", "\\1", param))
  ggplot(sud, aes_string("est", "forcats::fct_rev(par_value)",
    yend = "forcats::fct_rev(par_value)"
  )) +
    ggplot2::geom_segment(aes_string(
      x = "est - 1.96 * se",
      xend = "est + 1.96 * se"
    ), lwd = 0.5) +
    ggplot2::geom_segment(aes_string(
      x = "est - 0.67 * se",
      xend = "est + 0.67 * se"
    ), lwd = 1.25) +
    geom_point() +
    facet_wrap(~par_group, scales = "free") +
    theme_pbs() + guides(shape = FALSE, colour = FALSE) +
    labs(x = "Coefficient value (log space)", y = "Predictor value")
}
make_re_dat <- function(object) {
  re <- glmmTMB::ranef(object)
  plyr::ldply(re$cond, function(x) {
    sud <- as.data.frame(x)
    sud$par_value <- row.names(sud)
    row.names(sud) <- NULL
    sud
  }) %>%
    rename(par_group = .id) %>%
    rename(est = `(Intercept)`) %>%
    as_tibble() %>%
    mutate(loc_group = gsub("^([0-9]+)[ -]*([0-9a-zA-Z-]+)$", "\\2", par_value)) %>%
    mutate(loc_year = gsub("^([0-9]+)[ -]*[0-9a-zA-Z-]+$", "\\1", par_value))
}
  
make_re_plots <- function(object, re_names = c("locality")) {
  re <- make_re_dat(object)
  filter(re, par_group %in% re_names) %>% 
  ggplot(aes_string("est", "forcats::fct_rev(par_value)")) +
  geom_vline(xintercept = 0, lty = 2, alpha = 0.4) +
  geom_point(bg = "white") +
  facet_wrap(~par_group, scales = "free") +
  theme_pbs() + guides(shape = FALSE, colour = FALSE) +
  labs(x = "Random intercept value (log space)", y = "")
}
make_year_locality_plots <- function(object) {
  re <- make_re_dat(object)
  filter(re, par_group == "year_locality") %>%
    ggplot(aes_string("as.numeric(loc_year)", "est", group = "loc_group")) +
    geom_hline(yintercept = 0, lty = 2, alpha = 0.4) +
    geom_point(alpha = 0.7) +
    geom_line(alpha = 0.3) +
    facet_wrap(~loc_group) +
    scale_x_continuous(breaks = seq(1900, 3000, 10)) +
    theme_pbs() + guides(shape = FALSE, colour = FALSE) +
    labs(x = "Year", y = "Random intercept\n(interaction) value (log space)")
}
```

```{r cpue-coef-plot1, fig.asp=1.1, echo=FALSE}
toplot <- which(attr(
  model, "split_labels")$formula_version == "Full standardization")
for (i in seq_along(params$area_name)) {
  p <- make_fe_plots(model[[toplot[i]]]) + ggplot2::ggtitle(params$area_name[i])
  print(p)
}
```

```{r cpue-coef-plot2, fig.asp=0.8, echo=FALSE}
for (i in seq_along(params$area_name)) {
  p <- make_re_plots(model[[toplot[i]]],
    re_names = if (params$era == "modern") c("locality", "vessel") else "locality"
    ) + ggplot2::ggtitle(params$area_name[i])
  print(p)
}
```

```{r cpue-coef-plot3, fig.asp=length(params$area), echo=FALSE}
for (i in seq_along(params$area_name)) {
  p <- make_year_locality_plots(model[[toplot[i]]]) + ggplot2::ggtitle(params$area_name[i])
  print(p)
}
```

Now we will plot a comparison of the predictions at the base levels. The red line is the standardized version by one or more variables, the black solid line is a version with only a year predictor with the Tweedie observation model, the dashed line is the summed catch for the species divided by effort for all species each year. The ribbons indicate the 95% (Wald) confidence intervals.

```{r cpue-modern-predictions, fig.width=11, fig.asp=length(params$area)*0.2}
arith_cpue <- dfleet %>%
  bind_rows() %>% 
  group_by(area, year) %>%
  summarise(est = sum(spp_catch) / sum(hours_fished)) %>%
  mutate(model = "Combined") %>%
  group_by(area) %>%
  mutate(geo_mean = exp(mean(log(est)))) %>%
  mutate(est = est/geo_mean) %>%
  ungroup()

if (!params$skip_single_variable_models) {
  gfplot:::plot_cpue_predictions(predictions, "Combined", scale = TRUE) +
    geom_line(data = arith_cpue, aes(year, est),
      inherit.aes = FALSE, lty = 2) +
    scale_x_continuous(breaks = seq(1990, 2050, 5))
}
```

The index with the full standardization model for `r params$species_proper`:

```{r predictions-stand, fig.asp=length(params$area)*0.55}
predictions %>% 
  filter(formula_version %in% c("Unstandardized", "Full standardization")) %>% 
  gfplot:::plot_cpue_predictions("Combined", scale = TRUE) +
  geom_line(data = arith_cpue, aes(year, est),
    inherit.aes = FALSE, lty = 2) +
  scale_x_continuous(breaks = seq(1950, 2050, 5))
```

```{r predictions-stand-int, fig.asp=length(params$area)*0.55, echo=FALSE, eval=FALSE}
# predictions %>%
#   group_by(formula_version, model, area) %>%
#   mutate(geo_mean = exp(mean(log(est)))) %>%
#   mutate(upr = upr / geo_mean, lwr = lwr / geo_mean, est = est / geo_mean) %>%
#   ungroup() %>%
#   ggplot(aes(year, est, ymin = lwr, ymax = upr,
#     colour = formula_version, fill = formula_version)) + geom_line() +
#   geom_ribbon(alpha = 0.5) +
#   facet_wrap(~area, ncol = 1) +
#   ylab("CPUE (kg/hour) divided\nby geometric mean")
```

Session details:

```{r session, echo=FALSE}
si <- devtools::session_info(include_base = FALSE)
si$platform
packages <- dplyr::filter(si$packages, attached, !package %in% 
    c("usethis", "devtools", "graphics", "grDevices", 
      "utils", "datasets", "lintr", "methods", "stats", "bindrcpp",
      "testthat"))
dplyr::select(packages, package, loadedversion, date, source)
```

The specific version of the code used to generate this report can be found here:

```{r hash, cache=FALSE, echo=FALSE}
get_sha <- function(path) {
  suppressWarnings({
    sha <- tryCatch(system(paste0("git ls-remote git@github.com:", path, ".git"),
      intern = TRUE, ignore.stderr = TRUE), error = function(e) " ")
  })
  if (is.null(attr(sha, "status")))
    substr(sha[grepl("HEAD", sha)], 1, 7)
  else
    "xxxxxxx"
}
sha <- get_sha("pbs-assess/cpue-reports")
```

<`r paste0("https://github.com/pbs-assess/gfplot/tree/", sha)`>
